```{r}
library(rio)
library(dplyr)
library(stringr)
library(readr)
library(data.table)
library(tidyr)
```

# 1. ER
# 1.1. 2002

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-regionales-2002-oficina-nacional-de-procesos>

```{r}
ER2002 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2002_Gobernador_Vicegobernador.csv")
```

```{r}
d <- ER2002 %>% select(1:4,6,11:13) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataER02 <- ER2002 %>% select(1,10) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataER02 <- merge(dataER02, d, by = "UBIGEO")
dataER02 <- dataER02[,c(1,3,4,5,6,2,7,8)]
dataER02$TOTAL_VOTOS <- dataER02$VOTOS_OBTENIDOS + dataER02$VOTOS_BLANCOS + dataER02$VOTOS_NULOS
dataER02 <- dataER02[,-c(6)]
```

```{r}
# Calculations
dataER02$AUSENTISMO <- 1 - round(dataER02$TOTAL_VOTOS / dataER02$ELECTORES_HABILES,2)
dataER02$FreqBLANCOS <- round(dataER02$VOTOS_BLANCOS / dataER02$TOTAL_VOTOS,2)
dataER02$FreqNULOS <- round(dataER02$VOTOS_NULOS / dataER02$TOTAL_VOTOS,2)
dataER02$FreqINVALIDOS <- dataER02$FreqBLANCOS + dataER02$FreqNULOS

# Last steps
dataER02$YEAR = 2002
dataER02$TYPE = "REGIONALES"
```


## 1.2. 2006

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-de-gobernadores-y-vicegobernadores-regionales-2006-0>

```{r}
ER2006 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2006_Gobernador_Vicegobernador.csv")
```

```{r}
d <- ER2006 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataER06 <- ER2006 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataER06 <- merge(dataER06, d, by = "UBIGEO")
dataER06 <- dataER06[,c(1,3,4,5,6,2,7,8)]
dataER06$TOTAL_VOTOS <- dataER06$VOTOS_OBTENIDOS + dataER06$VOTOS_BLANCOS + dataER06$VOTOS_NULOS
dataER06 <- dataER06[,-c(6)]
```

```{r}
# Calculations
dataER06$AUSENTISMO <- 1 - round(dataER06$TOTAL_VOTOS / dataER06$ELECTORES_HABILES,2)
dataER06$FreqBLANCOS <- round(dataER06$VOTOS_BLANCOS / dataER06$TOTAL_VOTOS,2)
dataER06$FreqNULOS <- round(dataER06$VOTOS_NULOS / dataER06$TOTAL_VOTOS,2)
dataER06$FreqINVALIDOS <- dataER06$FreqBLANCOS + dataER06$FreqNULOS

# Last steps
dataER06$YEAR = 2006
dataER06$TYPE = "REGIONALES"
```


## 1.3. 2010

TOUGH BASTARD THIS ONE

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-de-gobernadores-y-vicegobernadores-regionales-2010-0>

```{r}
# Step 1: Read the file with proper encoding and delimiter
df <- read_delim("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2010_Gobernador_Vicegobernador.csv", delim = ";", locale = locale(encoding = "LATIN1"))

df <- df %>%
  mutate(across(everything(), ~ str_replace_all(., 
           "PARTIDO POPULAR CRISTIANO - PPC", 
           "PARTIDO POPULAR CRISTIANO - PPC - UNIDAD NACIONAL")))
```

```{r}
# Detect rows where the first column has the party name
mask <- grepl("^PARTIDO POPULAR CRISTIANO - PPC - UNIDAD NACIONAL", df[[1]])

# Shift those rows by adding 10 NAs at the beginning and removing the last 10 elements
df[mask, ] <- t(apply(df[mask, ], 1, function(row) {
  c(rep(NA, 10), row[1:(length(row) - 10)])
}))
```

```{r}
df <- df %>%
  mutate(
    VOTOS_OBTENIDOS = coalesce(VOTOS_OBTENIDOS, lead(VOTOS_OBTENIDOS)),
    ELECTORES_HABILES = coalesce(ELECTORES_HABILES, lead(ELECTORES_HABILES)),
    VOTOS_BLANCOS = coalesce(VOTOS_BLANCOS, lead(VOTOS_BLANCOS)),
    VOTOS_NULOS = coalesce(VOTOS_NULOS, lead(VOTOS_NULOS)),
    VOTOS_IMPUG = coalesce(VOTOS_IMPUG, lead(VOTOS_IMPUG))
  )

df <- df %>%
  filter(!is.na(UBIGEO))
```

```{r}
d <- df %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d$ELECTORES_HABILES <- as.numeric(d$ELECTORES_HABILES)
d$VOTOS_BLANCOS <- as.numeric(d$VOTOS_BLANCOS)
d$VOTOS_NULOS <- as.numeric(d$VOTOS_NULOS)

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataER10 <- df %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO))

dataER10$VOTOS_OBTENIDOS <- as.numeric(dataER10$VOTOS_OBTENIDOS)

dataER10 <- dataER10 %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataER10 <- merge(dataER10, d, by = "UBIGEO")
dataER10 <- dataER10[, c(1, 3, 4, 5, 6, 2, 7, 8)]
dataER10$TOTAL_VOTOS <- dataER10$VOTOS_OBTENIDOS + dataER10$VOTOS_BLANCOS + dataER10$VOTOS_NULOS
dataER10 <- dataER10[, -c(6)]
```

```{r}
# Calculations
dataER10$AUSENTISMO <- 1 - round(dataER10$TOTAL_VOTOS / dataER10$ELECTORES_HABILES, 2)
dataER10$FreqBLANCOS <- round(dataER10$VOTOS_BLANCOS / dataER10$TOTAL_VOTOS, 2)
dataER10$FreqNULOS <- round(dataER10$VOTOS_NULOS / dataER10$TOTAL_VOTOS, 2)
dataER10$FreqINVALIDOS <- dataER10$FreqBLANCOS + dataER10$FreqNULOS

# Last steps
dataER10$YEAR <- 2010
dataER10$TYPE <- "REGIONALES"
```


## 1.4. 2014

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-de-gobernadores-y-vicegobernadores-regionales-2014>

```{r}
ER2014 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2014_Gobernador_Vicegobernador.csv")
```

```{r}
d <- ER2014 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataER14 <- ER2014 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataER14 <- merge(dataER14, d, by = "UBIGEO")
dataER14 <- dataER14[,c(1,3,4,5,6,2,7,8)]
dataER14$TOTAL_VOTOS <- dataER14$VOTOS_OBTENIDOS + dataER14$VOTOS_BLANCOS + dataER14$VOTOS_NULOS
dataER14 <- dataER14[,-c(6)]
```

```{r}
# Calculations
dataER14$AUSENTISMO <- 1 - round(dataER14$TOTAL_VOTOS / dataER14$ELECTORES_HABILES, 2)
dataER14$FreqBLANCOS <- round(dataER14$VOTOS_BLANCOS / dataER14$TOTAL_VOTOS, 2)
dataER14$FreqNULOS <- round(dataER14$VOTOS_NULOS / dataER14$TOTAL_VOTOS, 2)
dataER14$FreqINVALIDOS <- dataER14$FreqBLANCOS + dataER14$FreqNULOS

# Last steps
dataER14$YEAR = 2014
dataER14$TYPE = "REGIONALES"
```


## 1.5. 2018

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-de-gobernadores-y-vicegobernadores-regionales-2018-1>

```{r}
ER2018 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2018_Regional_total.csv")
```

```{r}
d <- ER2018 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataER18 <- ER2018 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataER18 <- merge(dataER18, d, by = "UBIGEO")
dataER18 <- dataER18[,c(1,3,4,5,6,2,7,8)]
dataER18$TOTAL_VOTOS <- dataER18$VOTOS_OBTENIDOS + dataER18$VOTOS_BLANCOS + dataER18$VOTOS_NULOS
dataER18 <- dataER18[,-c(6)]
```

```{r}
# Calculations
dataER18$AUSENTISMO <- 1 - round(dataER18$TOTAL_VOTOS / dataER18$ELECTORES_HABILES, 2)
dataER18$FreqBLANCOS <- round(dataER18$VOTOS_BLANCOS / dataER18$TOTAL_VOTOS, 2)
dataER18$FreqNULOS <- round(dataER18$VOTOS_NULOS / dataER18$TOTAL_VOTOS, 2)
dataER18$FreqINVALIDOS <- dataER18$FreqBLANCOS + dataER18$FreqNULOS

# Last steps
dataER18$YEAR = 2018
dataER18$TYPE = "REGIONALES"
```

## 1.6. 2022

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-de-gobernadores-y-vicegobernadores-regionales-2022>

```{r}
ER2022 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2022_Gobernador_Vicegobernador.csv")
```

```{r}
d <- ER2022 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataER22 <- ER2022 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataER22 <- merge(dataER22, d, by = "UBIGEO")
dataER22 <- dataER22[,c(1,3,4,5,6,2,7,8)]
dataER22$TOTAL_VOTOS <- dataER22$VOTOS_OBTENIDOS + dataER22$VOTOS_BLANCOS + dataER22$VOTOS_NULOS
dataER22 <- dataER22[,-c(6)]
```

```{r}
# Calculations
dataER22$AUSENTISMO <- 1 - round(dataER22$TOTAL_VOTOS / dataER22$ELECTORES_HABILES, 2)
dataER22$FreqBLANCOS <- round(dataER22$VOTOS_BLANCOS / dataER22$TOTAL_VOTOS, 2)
dataER22$FreqNULOS <- round(dataER22$VOTOS_NULOS / dataER22$TOTAL_VOTOS, 2)
dataER22$FreqINVALIDOS <- dataER22$FreqBLANCOS + dataER22$FreqNULOS

# Last steps
dataER22$YEAR = 2022
dataER22$TYPE = "REGIONALES"
```

```{r}
# Append
regionales <- rbind(dataER02, dataER06)
regionales <- rbind(regionales, dataER10)
regionales <- rbind(regionales, dataER14)
regionales <- rbind(regionales, dataER18)
regionales <- rbind(regionales, dataER22)
```



# 2. EP
## 2.1. 2002

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-municipales-provinciales-2002-oficina-nacional-de>

Lima province is missing here. Numbers match if included.

```{r}
EP2002 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2002_Municipal_Provincial.csv")
```

```{r}
d <- EP2002 %>% select(1:4,6,11:13) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataEP02 <- EP2002 %>% select(1,10) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataEP02 <- merge(dataEP02, d, by = "UBIGEO")
dataEP02 <- dataEP02[,c(1,3,4,5,6,2,7,8)]
dataEP02$TOTAL_VOTOS <- dataEP02$VOTOS_OBTENIDOS + dataEP02$VOTOS_BLANCOS + dataEP02$VOTOS_NULOS
dataEP02 <- dataEP02[,-c(6)]
```

```{r}
# Calculations
dataEP02$AUSENTISMO <- 1 - round(dataEP02$TOTAL_VOTOS / dataEP02$ELECTORES_HABILES, 2)
dataEP02$FreqBLANCOS <- round(dataEP02$VOTOS_BLANCOS / dataEP02$TOTAL_VOTOS, 2)
dataEP02$FreqNULOS <- round(dataEP02$VOTOS_NULOS / dataEP02$TOTAL_VOTOS, 2)
dataEP02$FreqINVALIDOS <- dataEP02$FreqBLANCOS + dataEP02$FreqNULOS

# Last steps
dataEP02$YEAR = 2002
dataEP02$TYPE = "PROVINCIALES"
```


## 2.2. 2006

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-municipales-provinciales-2006-oficina-nacional-de>

All data is here.

```{r}
EP2006 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2006_Municipal_Provincial.csv")
```

```{r}
d <- EP2006 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataEP06 <- EP2006 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataEP06 <- merge(dataEP06, d, by = "UBIGEO")
dataEP06 <- dataEP06[,c(1,3,4,5,6,2,7,8)]
dataEP06$TOTAL_VOTOS <- dataEP06$VOTOS_OBTENIDOS + dataEP06$VOTOS_BLANCOS + dataEP06$VOTOS_NULOS
dataEP06 <- dataEP06[,-c(6)]
```

```{r}
# Calculations
dataEP06$AUSENTISMO <- 1 - round(dataEP06$TOTAL_VOTOS / dataEP06$ELECTORES_HABILES, 2)
dataEP06$FreqBLANCOS <- round(dataEP06$VOTOS_BLANCOS / dataEP06$TOTAL_VOTOS, 2)
dataEP06$FreqNULOS <- round(dataEP06$VOTOS_NULOS / dataEP06$TOTAL_VOTOS, 2)
dataEP06$FreqINVALIDOS <- dataEP06$FreqBLANCOS + dataEP06$FreqNULOS

# Last steps
dataEP06$YEAR = 2006
dataEP06$TYPE = "PROVINCIALES"
```


## 2.3. 2010

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-municipales-provinciales-2010-oficina-nacional-de-0>

All data is here.

```{r}
# Step 1: Read the file with proper encoding and delimiter
df <- read_delim("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2010_Municipal_Provincial.csv", delim = ";", locale = locale(encoding = "LATIN1"))

df <- df %>%
  mutate(across(everything(), ~ str_replace_all(., 
           "PARTIDO POPULAR CRISTIANO - PPC", 
           "PARTIDO POPULAR CRISTIANO - PPC - UNIDAD NACIONAL")))

```

```{r}
# Detect rows where the first column has the party name
mask <- grepl("^PARTIDO POPULAR CRISTIANO - PPC - UNIDAD NACIONAL", df[[1]])

# Shift those rows by adding 10 NAs at the beginning and removing the last 10 elements
df[mask, ] <- t(apply(df[mask, ], 1, function(row) {
  c(rep(NA, 10), row[1:(length(row) - 10)])
}))
```

```{r}
df <- df %>%
  mutate(
    VOTOS_OBTENIDOS = coalesce(VOTOS_OBTENIDOS, lead(VOTOS_OBTENIDOS)),
    ELECTORES_HABILES = coalesce(ELECTORES_HABILES, lead(ELECTORES_HABILES)),
    VOTOS_BLANCOS = coalesce(VOTOS_BLANCOS, lead(VOTOS_BLANCOS)),
    VOTOS_NULOS = coalesce(VOTOS_NULOS, lead(VOTOS_NULOS)),
    VOTOS_IMPUG = coalesce(VOTOS_IMPUG, lead(VOTOS_IMPUG))
  )

df <- df %>%
  filter(!is.na(UBIGEO))
```

```{r}
d <- df %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d$ELECTORES_HABILES <- as.numeric(d$ELECTORES_HABILES)
d$VOTOS_BLANCOS <- as.numeric(d$VOTOS_BLANCOS)
d$VOTOS_NULOS <- as.numeric(d$VOTOS_NULOS)

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataEP10 <- df %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO))

dataEP10$VOTOS_OBTENIDOS <- as.numeric(dataEP10$VOTOS_OBTENIDOS)

dataEP10 <- dataEP10 %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataEP10 <- merge(dataEP10, d, by = "UBIGEO")
dataEP10 <- dataEP10[, c(1, 3, 4, 5, 6, 2, 7, 8)]
dataEP10$TOTAL_VOTOS <- dataEP10$VOTOS_OBTENIDOS + dataEP10$VOTOS_BLANCOS + dataEP10$VOTOS_NULOS
dataEP10 <- dataEP10[, -c(6)]
```

```{r}
# Calculations
dataEP10$AUSENTISMO <- 1 - round(dataEP10$TOTAL_VOTOS / dataEP10$ELECTORES_HABILES, 2)
dataEP10$FreqBLANCOS <- round(dataEP10$VOTOS_BLANCOS / dataEP10$TOTAL_VOTOS, 2)
dataEP10$FreqNULOS <- round(dataEP10$VOTOS_NULOS / dataEP10$TOTAL_VOTOS, 2)
dataEP10$FreqINVALIDOS <- dataEP10$FreqBLANCOS + dataEP10$FreqNULOS

# Last steps
dataEP10$YEAR = 2010
dataEP10$TYPE = "PROVINCIALES"
```


## 2.4. 2014

All data is here.

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-municipales-provinciales-2014-oficina-nacional-de>

```{r}
EP2014 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2014_Municipal_Provincial.csv")
```

```{r}
d <- EP2014 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataEP14 <- EP2014 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataEP14 <- merge(dataEP14, d, by = "UBIGEO")
dataEP14 <- dataEP14[,c(1,3,4,5,6,2,7,8)]
dataEP14$TOTAL_VOTOS <- dataEP14$VOTOS_OBTENIDOS + dataEP14$VOTOS_BLANCOS + dataEP14$VOTOS_NULOS
dataEP14 <- dataEP14[,-c(6)]
```

```{r}
# Calculations
dataEP14$AUSENTISMO <- 1 - round(dataEP14$TOTAL_VOTOS / dataEP14$ELECTORES_HABILES, 2)
dataEP14$FreqBLANCOS <- round(dataEP14$VOTOS_BLANCOS / dataEP14$TOTAL_VOTOS, 2)
dataEP14$FreqNULOS <- round(dataEP14$VOTOS_NULOS / dataEP14$TOTAL_VOTOS, 2)
dataEP14$FreqINVALIDOS <- dataEP14$FreqBLANCOS + dataEP14$FreqNULOS

# Last steps
dataEP14$YEAR = 2014
dataEP14$TYPE = "PROVINCIALES"
```


## 2.5. 2018

All data here

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-municipales-provinciales-2018-oficina-nacional-de-1>

```{r}
EP2018 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2018_Municipal_Provincial_total.csv")
```

```{r}
d <- EP2018 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataEP18 <- EP2018 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataEP18 <- merge(dataEP18, d, by = "UBIGEO")
dataEP18 <- dataEP18[,c(1,3,4,5,6,2,7,8)]
dataEP18$TOTAL_VOTOS <- dataEP18$VOTOS_OBTENIDOS + dataEP18$VOTOS_BLANCOS + dataEP18$VOTOS_NULOS
dataEP18 <- dataEP18[,-c(6)]
```

```{r}
# Calculations
dataEP18$AUSENTISMO <- 1 - round(dataEP18$TOTAL_VOTOS / dataEP18$ELECTORES_HABILES, 2)
dataEP18$FreqBLANCOS <- round(dataEP18$VOTOS_BLANCOS / dataEP18$TOTAL_VOTOS, 2)
dataEP18$FreqNULOS <- round(dataEP18$VOTOS_NULOS / dataEP18$TOTAL_VOTOS, 2)
dataEP18$FreqINVALIDOS <- dataEP18$FreqBLANCOS + dataEP18$FreqNULOS

# Last steps
dataEP18$YEAR = 2018
dataEP18$TYPE = "PROVINCIALES"
```


## 2.6. 2022

Link: <https://datosabiertos.gob.pe/dataset/resultados-por-mesa-de-las-elecciones-municipales-provinciales-2022-oficina-nacional-de>

```{r}
EP2022 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2022_Municipal_Provincial.csv")
```

```{r}
d <- EP2022 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataEP22 <- EP2022 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataEP22 <- merge(dataEP22, d, by = "UBIGEO")
dataEP22 <- dataEP22[,c(1,3,4,5,6,2,7,8)]
dataEP22$TOTAL_VOTOS <- dataEP22$VOTOS_OBTENIDOS + dataEP22$VOTOS_BLANCOS + dataEP22$VOTOS_NULOS
dataEP22 <- dataEP22[,-c(6)]
```

```{r}
# Calculations
dataEP22$AUSENTISMO <- 1 - round(dataEP22$TOTAL_VOTOS / dataEP22$ELECTORES_HABILES, 2)
dataEP22$FreqBLANCOS <- round(dataEP22$VOTOS_BLANCOS / dataEP22$TOTAL_VOTOS, 2)
dataEP22$FreqNULOS <- round(dataEP22$VOTOS_NULOS / dataEP22$TOTAL_VOTOS, 2)
dataEP22$FreqINVALIDOS <- dataEP22$FreqBLANCOS + dataEP22$FreqNULOS

# Last steps
dataEP22$YEAR = 2022
dataEP22$TYPE = "PROVINCIALES"
```


Solve 2002 with Lima province missing. We have to add 2002 district elections.

# 3. Districts
## 3.1. 2002

Lima missing again. Numbers match if included.

```{r}
ED2002 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2002_Municipal_Distrital.csv")
```

```{r}
d <- ED2002 %>% select(1:4,6,11:13) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataED02 <- ED2002 %>% select(1,10) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataED02 <- merge(dataED02, d, by = "UBIGEO")
dataED02 <- dataED02[,c(1,3,4,5,6,2,7,8)]
dataED02$TOTAL_VOTOS <- dataED02$VOTOS_OBTENIDOS + dataED02$VOTOS_BLANCOS + dataED02$VOTOS_NULOS
dataED02 <- dataED02[,-c(6)]
```

```{r}
# Calculations
dataED02$AUSENTISMO <- 1 - round(dataED02$TOTAL_VOTOS / dataED02$ELECTORES_HABILES, 2)
dataED02$FreqBLANCOS <- round(dataED02$VOTOS_BLANCOS / dataED02$TOTAL_VOTOS, 2)
dataED02$FreqNULOS <- round(dataED02$VOTOS_NULOS / dataED02$TOTAL_VOTOS, 2)
dataED02$FreqINVALIDOS <- dataED02$FreqBLANCOS + dataED02$FreqNULOS

# Last steps
dataED02$YEAR = 2002
dataED02$TYPE = "DISTRITALES"
```


## 3.2. 2006

All data here. 

```{r}
ED2006 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2006_Municipal_Distrital.csv")
```

```{r}
d <- ED2006 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataED06 <- ED2006 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataED06 <- merge(dataED06, d, by = "UBIGEO")
dataED06 <- dataED06[,c(1,3,4,5,6,2,7,8)]
dataED06$TOTAL_VOTOS <- dataED06$VOTOS_OBTENIDOS + dataED06$VOTOS_BLANCOS + dataED06$VOTOS_NULOS
dataED06 <- dataED06[,-c(6)]
```

```{r}
# Calculations
dataED06$AUSENTISMO <- 1 - round(dataED06$TOTAL_VOTOS / dataED06$ELECTORES_HABILES, 2)
dataED06$FreqBLANCOS <- round(dataED06$VOTOS_BLANCOS / dataED06$TOTAL_VOTOS, 2)
dataED06$FreqNULOS <- round(dataED06$VOTOS_NULOS / dataED06$TOTAL_VOTOS, 2)
dataED06$FreqINVALIDOS <- dataED06$FreqBLANCOS + dataED06$FreqNULOS

# Last steps
dataED06$YEAR = 2006
dataED06$TYPE = "DISTRITALES"
```


## 3.3. 2010

All data here.

```{r}
# Step 1: Read the file with proper encoding and delimiter
df <- read_delim("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2010_Municipal_Distrital.csv", delim = ";", locale = locale(encoding = "LATIN1"))

df <- df %>%
  mutate(across(everything(), ~ str_replace_all(., 
           "PARTIDO POPULAR CRISTIANO - PPC", 
           "PARTIDO POPULAR CRISTIANO - PPC - UNIDAD NACIONAL")))

```

```{r}
# Detect rows where the first column has the party name
mask <- grepl("^PARTIDO POPULAR CRISTIANO - PPC - UNIDAD NACIONAL", df[[1]])

# Shift those rows by adding 10 NAs at the beginning and removing the last 10 elements
df[mask, ] <- t(apply(df[mask, ], 1, function(row) {
  c(rep(NA, 10), row[1:(length(row) - 10)])
}))
```

```{r}
df <- df %>%
  mutate(
    VOTOS_OBTENIDOS = coalesce(VOTOS_OBTENIDOS, lead(VOTOS_OBTENIDOS)),
    ELECTORES_HABILES = coalesce(ELECTORES_HABILES, lead(ELECTORES_HABILES)),
    VOTOS_BLANCOS = coalesce(VOTOS_BLANCOS, lead(VOTOS_BLANCOS)),
    VOTOS_NULOS = coalesce(VOTOS_NULOS, lead(VOTOS_NULOS)),
    VOTOS_IMPUG = coalesce(VOTOS_IMPUG, lead(VOTOS_IMPUG))
  )

df <- df %>%
  filter(!is.na(UBIGEO))
```

```{r}
d <- df %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d$ELECTORES_HABILES <- as.numeric(d$ELECTORES_HABILES)
d$VOTOS_BLANCOS <- as.numeric(d$VOTOS_BLANCOS)
d$VOTOS_NULOS <- as.numeric(d$VOTOS_NULOS)

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataED10 <- df %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO))

dataED10$VOTOS_OBTENIDOS <- as.numeric(dataED10$VOTOS_OBTENIDOS)

dataED10 <- dataED10 %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataED10 <- merge(dataED10, d, by = "UBIGEO")
dataED10 <- dataED10[, c(1, 3, 4, 5, 6, 2, 7, 8)]
dataED10$TOTAL_VOTOS <- dataED10$VOTOS_OBTENIDOS + dataED10$VOTOS_BLANCOS + dataED10$VOTOS_NULOS
dataED10 <- dataED10[, -c(6)]
```

```{r}
# Calculations
dataED10$AUSENTISMO <- 1 - round(dataED10$TOTAL_VOTOS / dataED10$ELECTORES_HABILES, 2)
dataED10$FreqBLANCOS <- round(dataED10$VOTOS_BLANCOS / dataED10$TOTAL_VOTOS, 2)
dataED10$FreqNULOS <- round(dataED10$VOTOS_NULOS / dataED10$TOTAL_VOTOS, 2)
dataED10$FreqINVALIDOS <- dataED10$FreqBLANCOS + dataED10$FreqNULOS

# Last steps
dataED10$YEAR = 2010
dataED10$TYPE = "DISTRITALES"
```


## 3.4. 2014

All data here.

```{r}
ED2014 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2014_Municipal_Distrital.csv")
```

```{r}
d <- ED2014 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataED14 <- ED2014 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataED14 <- merge(dataED14, d, by = "UBIGEO")
dataED14 <- dataED14[, c(1, 3, 4, 5, 6, 2, 7, 8)]
dataED14$TOTAL_VOTOS <- dataED14$VOTOS_OBTENIDOS + dataED14$VOTOS_BLANCOS + dataED14$VOTOS_NULOS
dataED14 <- dataED14[, -c(6)]
```

```{r}
# Calculations
dataED14$AUSENTISMO <- 1 - round(dataED14$TOTAL_VOTOS / dataED14$ELECTORES_HABILES, 2)
dataED14$FreqBLANCOS <- round(dataED14$VOTOS_BLANCOS / dataED14$TOTAL_VOTOS, 2)
dataED14$FreqNULOS <- round(dataED14$VOTOS_NULOS / dataED14$TOTAL_VOTOS, 2)
dataED14$FreqINVALIDOS <- dataED14$FreqBLANCOS + dataED14$FreqNULOS

# Last steps
dataED14$YEAR = 2014
dataED14$TYPE = "DISTRITALES"
```


## 3.5. 2018

All here.

```{r}
ED2018 <- import("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2018_Municipal_Distrital_total.csv")
```

```{r}
d <- ED2018 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataED18 <- ED2018 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataED18 <- merge(dataED18, d, by = "UBIGEO")
dataED18 <- dataED18[, c(1, 3, 4, 5, 6, 2, 7, 8)]
dataED18$TOTAL_VOTOS <- dataED18$VOTOS_OBTENIDOS + dataED18$VOTOS_BLANCOS + dataED18$VOTOS_NULOS
dataED18 <- dataED18[, -c(6)]
```

```{r}
# Calculations
dataED18$AUSENTISMO <- 1 - round(dataED18$TOTAL_VOTOS / dataED18$ELECTORES_HABILES, 2)
dataED18$FreqBLANCOS <- round(dataED18$VOTOS_BLANCOS / dataED18$TOTAL_VOTOS, 2)
dataED18$FreqNULOS <- round(dataED18$VOTOS_NULOS / dataED18$TOTAL_VOTOS, 2)
dataED18$FreqINVALIDOS <- dataED18$FreqBLANCOS + dataED18$FreqNULOS

# Last steps
dataED18$YEAR = 2018
dataED18$TYPE = "DISTRITALES"
```


## 3.6. 2022

All data here. 

```{r}
# Rio not working properly.
ED2022 <- read_delim("/Users/alexander/Documents/boyco-benites/data/raw/subnacional/ERM2022_Municipal_Distrital.csv", delim = ";", locale = locale(encoding = "LATIN1"))
```

```{r}
d <- ED2022 %>% select(1:4,6,13:15) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  distinct() # Unique cases

d <- d %>% select(-MESA) %>% 
  group_by(UBIGEO,DEPARTAMENTO, PROVINCIA, DISTRITO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
dataED22 <- ED2022 %>% select(1,12) %>% # Adding zeros:
  mutate(UBIGEO = ifelse(nchar(UBIGEO) == 5, str_pad(UBIGEO, 6, pad = "0"), UBIGEO)) %>%
  group_by(UBIGEO) %>%  # Group and sum by district
  summarise(across(where(is.numeric), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
```

```{r}
# Calculating total votes per district
dataED22 <- merge(dataED22, d, by = "UBIGEO")
dataED22 <- dataED22[, c(1, 3, 4, 5, 6, 2, 7, 8)]
dataED22$TOTAL_VOTOS <- dataED22$VOTOS_OBTENIDOS + dataED22$VOTOS_BLANCOS + dataED22$VOTOS_NULOS
dataED22 <- dataED22[, -c(6)]
```

```{r}
# Calculations
dataED22$AUSENTISMO <- 1 - round(dataED22$TOTAL_VOTOS / dataED22$ELECTORES_HABILES, 2)
dataED22$FreqBLANCOS <- round(dataED22$VOTOS_BLANCOS / dataED22$TOTAL_VOTOS, 2)
dataED22$FreqNULOS <- round(dataED22$VOTOS_NULOS / dataED22$TOTAL_VOTOS, 2)
dataED22$FreqINVALIDOS <- dataED22$FreqBLANCOS + dataED22$FreqNULOS

# Last steps
dataED22$YEAR = 2022
dataED22$TYPE = "DISTRITALES"
```


## Handling inconsistencies

To be solved: we have to include results for Lima Province Municipality in 2002, and District Municipality elections in 2002. Missing in original data.

We start with district elections from 2002. From Infogob:

```{r}
## Add Lima
ubi2002 <- import("/Users/alexander/Documents/boyco-benites/data/raw/prob/ubi2002.xlsx")
e2002 <- import("/Users/alexander/Documents/boyco-benites/data/raw/prob/munidis2002.xlsx")

e2002 <- e2002[e2002$Provincia == "LIMA",]
ubi2002 <- ubi2002[ubi2002$PROVINCIA == "LIMA",]

d <- e2002 %>% select(1,2,3,8,10)
e2002 <- e2002 %>% select(1,2,3,4,6) %>% distinct()
ubi2002 <- ubi2002 %>% select(1,4)

names(e2002) <- c("DEPARTAMENTO","PROVINCIA","DISTRITO","ELECTORES_HABILES","TOTAL_VOTOS")
e2002 <- merge(e2002, ubi2002, by = "DISTRITO")

d <- d %>%
  filter(`Organización Política` %in% c("VOTOS EN BLANCO", "VOTOS NULOS")) %>%
  pivot_wider(
    names_from = `Organización Política`,
    values_from = Votos
  ) %>% select(3:5)

names(d) <- c("DISTRITO","VOTOS_BLANCOS","VOTOS_NULOS")

e2002 <- merge(e2002, d, by = "DISTRITO")
e2002 <- e2002[,c(6,2,3,1,4,7,8,5)]
```

```{r}
# Calculations
e2002$AUSENTISMO <- 1 - round(e2002$TOTAL_VOTOS / e2002$ELECTORES_HABILES, 2)
e2002$FreqBLANCOS <- round(e2002$VOTOS_BLANCOS / e2002$TOTAL_VOTOS, 2)
e2002$FreqNULOS <- round(e2002$VOTOS_NULOS / e2002$TOTAL_VOTOS, 2)
e2002$FreqINVALIDOS <- e2002$FreqBLANCOS + e2002$FreqNULOS

# Last steps
e2002$YEAR = 2002
e2002$TYPE = "DISTRITALES"

# Stack
dataED02 <- rbind(dataED02, e2002)
```

Now Province:

15,298,237 for the election

```{r}
sum(dataEP02$ELECTORES_HABILES)
```

## Stack

```{r}
provinciales <- rbind(dataEP02, dataEP06)
provinciales <- rbind(provinciales, dataEP10)
provinciales <- rbind(provinciales, dataEP14)
provinciales <- rbind(provinciales, dataEP18)
provinciales <- rbind(provinciales, dataEP22)
```

```{r}
distritales <- rbind(dataED02, dataED06)
distritales <- rbind(distritales, dataED10)
distritales <- rbind(distritales, dataED14)
distritales <- rbind(distritales, dataED18)
distritales <- rbind(distritales, dataED22)
```

